|

### 2021-12-08 01-Chp-Background.Rmd
Deal with abbreviations, fix cross-links and some other small edits

Vascular dementia (VaD) is the second largest underlying pathology of dementia, accounting for ~10% of cases. Vascular dementia is caused by a range of cerebrovascular disorders, and as a result, presentation  of symptoms can vary widely.[@iadecola2013] Similarly, due to the varied underlying  pathophysiology, vascular dementia can onset either quite rapidly following a cerebrovascular event such as a stroke or over a long time-frame due to a series of small infarcts.[@venkat2015] Vascular dementia regularly co-occurs in patients with Alzheimer's disease.[@iadecola2013] This presentation is described as "mixed" dementia,[@custodio2017] and occurs in approximately 25% of cases.[@burns2009]

The remaining 10-30% of cases are caused by other dementia subtypes (e.g. Lewy body dementia, frontotemporal dementia) or by progression of other neurological diseases (e.g. Parkinson's disease).[@burns2009]

@@

By far the largest source of evidence on the relationship between blood lipids and dementia outcomes comes from observational designs. Several studies have examined the relationships between concentrations of serum lipids (total cholesterol (TC), low density lipoprotein cholesterol (LDL-c), high density lipoprotein cholesterol (HDL-c) and triglycerides) and both Alzheimer's disease and vascular dementia and reported extremely varied results. In some studies, a high serum cholesterol concentration has been found to be associated with an increase in susceptibility to AD,[@kivipelto2002; @kivipelto2005; @schilling2017; @solomon2009; @whitmer2005] however others have shown no association,[@li2005; @mainous2005; @mielke2010; @tan2003] or a reduced susceptibility [@mielke2005; @reitz2004]. With regard to vascular dementia, decreased levels of HDL-c appear to be associated with increased risk [@reitz2004], while for LDL-c, studies have reported both positive and negative associations.[@reitz2004; @moroney1999]

@@

However, the role of preprints in evidence syntheses is less well defined. Defined by the Committee on Publication Ethics (COPE) as 'scholarly manuscript[s] posted by the author(s) in an openly accessible platform, usually before or in parallel with the peer review process'[@committeeonpublicationethicscope2018], preprints serve several purposes. They are used to establish primacy when submitting to a journal where the peer-review process may take several months.[@vale2016] In addition, the allow for the rapid disseminate research findings, as occurred during the COVID-19 pandemic.[@fraser2020preprinting] Finally, they are a source of information on manuscripts that may not have been accepted elsewhere, helping to combat publication bias or the "file-drawer" effect.[@rosenthal1979]

@@

### Contributions to the scientific literature

>>>>

Vascular dementia (VaD) is the second largest underlying pathology of dementia, accounting for ~10-25% of cases. Vascular dementia is caused by a range of cerebrovascular disorders, and as a result, presentation  of symptoms can vary widely.[@iadecola2013] Similarly, due to the varied underlying  pathophysiology, vascular dementia can onset either quite rapidly following a cerebrovascular event such as a stroke or over a long time-frame due to a series of small infarcts.[@venkat2015] Vascular dementia regularly co-occurs in patients with Alzheimer's disease.[@iadecola2013] This presentation is described as "mixed" dementia,[@custodio2017] and occurs in approximately 25% of cases.[@burns2009]

The remaining 10-30% of cases are caused by other dementia subtypes (e.g. Lewy body or frontotemporal dementias) or other neurological diseases (e.g. Parkinson's disease).[@burns2009]

@@

By far the largest source of evidence on the relationship between blood lipids and dementia outcomes comes from observational designs. Several studies have examined the relationships between concentrations of serum lipids (total cholesterol (TC), low density lipoprotein cholesterol (LDL-c), high density lipoprotein cholesterol (HDL-c) and triglycerides) <!-- TODO These abbreviations have already been defined above --> and both Alzheimer's disease and vascular dementia and reported extremely varied results. In some studies, a high serum cholesterol concentration has been found to be associated with an increase in susceptibility to AD,[@kivipelto2002; @kivipelto2005; @schilling2017; @solomon2009; @whitmer2005] however others have shown no association,[@li2005; @mainous2005; @mielke2010; @tan2003] or a reduced susceptibility [@mielke2005; @reitz2004]. With regard to vascular dementia, decreased levels of HDL-c appear to be associated with increased risk [@reitz2004], while for LDL-c, studies have reported both positive and negative associations.[@reitz2004; @moroney1999]

@@

However, the role of preprints in evidence syntheses is less well defined. Defined by the Committee on Publication Ethics as 'scholarly manuscript[s] posted by the author(s) in an openly accessible platform, usually before or in parallel with the peer review process',[@committeeonpublicationethicscope2018] preprints serve several purposes. They are used to establish primacy when submitting to a journal where the peer-review process may take several months.[@vale2016] In addition, the allow for the rapid disseminate research findings, as occurred during the COVID-19 pandemic.[@fraser2020preprinting] Finally, they are a source of information on manuscripts that may not have been accepted elsewhere, helping to combat publication bias or the "file-drawer" effect.[@rosenthal1979]

@@

### Contributions to the scientific literature {#literature-contributions}

### 2021-12-08 02-Chp-Systematic-Review-Tools.Rmd
Deal with abbreviations, fix cross-links and some other small edits

However, an Application Programming Interface (API) for the medRxiv and bioRxiv repositories was made public in early 2020 by the institution responsible for managing these preprint repositories, the Cold Springs Harbor Laboratory. This allowed for newer versions of the `medrxivr` package to engage in active "fault prevention" and provide a more robust interface to the data by removing the reliance of web-scraping.[@laprie1992]

@@

```{r, eval = FALSE}
# Get a copy of the database from the live medRxiv API endpoint
mx_data <- mx_api_content()
```

@@

```{r, include=FALSE}
med_res <-
  # Use snapshot from July 2021
  medrxivr::mx_search(medrxivr::mx_snapshot("ccedfb8a44304b9fba4e3ba518a8ce4ed2294770"),
                      "*",
                      from_date = "2019-07-01",
                      to_date = "2019-08-01") %>%
mutate(pub_ind = ifelse(published == "NA", 0, 1)) %>%
  group_by(pub_ind) %>%
  count()

@@

Secondly, using `medrxivr`, an analysis of the publication rate for medRxiv preprints was performed (see Appendix \@ref(appendix-sys-rev-tools)). `r published` of the `r publication_n` records `r publish_percent` posted on medRxiv in July 2019 were published by 30th July 2021 (i.e. allowing for a two-year lag between preprint posting and publication). This finding agrees with previous work demonstrating that two-thirds of bioRxiv preprints are published in a peer-reviewed journal within two years of posting.[@abdill2019popularity] This analysis further illustrates that a non-insignificant number of preprints are never formally published but remain accessible as preprints.

>>>>

However, an application programming interface (API) for the medRxiv and bioRxiv repositories was made public in early 2020 by the institution responsible for managing these preprint repositories, the Cold Springs Harbor Laboratory. This allowed for newer versions of the `medrxivr` package to engage in active "fault prevention" and provide a more robust interface to the data by removing the reliance of web-scraping.[@laprie1992]

@@

```{r, eval = FALSE, class.source = "bg-danger"}
# Get a copy of the database from the live medRxiv API endpoint
mx_data <- mx_api_content()
```

@@

```{r, include=FALSE}
med_res <-
  # Use snapshot from July 2021
  medrxivr::mx_search(
    medrxivr::mx_snapshot("ccedfb8a44304b9fba4e3ba518a8ce4ed2294770"),
    query = "*",
    from_date = "2019-07-01",
    to_date = "2019-08-01"
  ) %>%
  mutate(pub_ind = ifelse(published == "NA", 0, 1)) %>%
  group_by(pub_ind) %>%
  count()

@@

Secondly, using `medrxivr`, an analysis of the publication rate for medRxiv preprints was performed (see Appendix \@ref(appendix-tools-pub)). `r published` of the `r publication_n` records `r publish_percent` posted on medRxiv in July 2019 were published by 30th July 2021 (i.e. allowing for a two-year lag between preprint posting and publication). This finding agrees with previous work demonstrating that two-thirds of bioRxiv preprints are published in a peer-reviewed journal within two years of posting.[@abdill2019popularity] This analysis further illustrates that a subtantial number of preprints are never formally published but remain accessible as preprints.

### 2021-12-08 03-Chp-Systematic-Review-Methods.Rmd
Fix broken links to appendices
Fix broken cross-links
Deal with abbreviations, fix cross-links and some other small edits

A pre-specified protocol for this analysis was registered on the Open Science Framework platform and is available for inspection.[@mcguinnessluke2020]

@@

Rater reliability was assessed using Gwet's agreement coefficient (AC1).[@gwet2008] This measure was chosen over other methods such as percent agreement (number of agreements divided by total number of assessments), as it accounts for chance agreement between reviewers but does not suffer from bias due to severely imbalanced marginal totals in the same way that Cohen's $kappa$ value does. [@cohen1960: @gwet2008; @wongpakaran2013] Given the small number of included studies in this review as a proportion of the total number screened, this is a useful characteristic.

@@

<!----------------------------------------------------------------------------->
(ref:gwet-caption) Suggested ranges to aid in interpretation of Gwet's AC1 inter-rater reliability metric

@@

This additional appraisal marks a departure from the registered protocol, as there was initially no intention to try to examine the risk of bias due to missing evidence. This is largely because the tool did not exist when the protocol was registered.

@@

#### Random-effects meta-analysis

@@

Several of the included studies presented data on multiple categories of lipid levels, but provided an overall effect estimate based on a comparison of only two of these categories (e.g. for example, highest vs lowest quartile). <!-- TODO Question: if I look at highest vs lowest, and also at top dose vs lowest dose, am I double counting results? --> While this allows for easy interpretation of the resulting effect estimate, it ignores any potential non-linear relationships between the exposure and outcome, in addition to discarding useful information contained in the interim groups. In order to address this limitation, I performed a dose-response meta-analysis in those studies reporting more than two categories of lipid levels. This marks a departure from the published protocol, as I had not expected the number of studies reporting dementia risk across lipid sub-categories to be sufficient for a dose response analysis.

>>>>

A pre-specified protocol for this analysis was registered on the Open Science Framework platform and is available for inspection.[@mcguinnessluke2020] Deviations from the protocol are summarised in Appendix \@ref(appendix-sys-rev-deviations).

@@

Rater reliability was assessed using Gwet's agreement coefficient (AC1).[@gwet2008] This measure was chosen over other methods such as percent agreement (number of agreements divided by total number of assessments), as it accounts for chance agreement between reviewers but does not suffer from bias due to severely imbalanced marginal totals in the same way that Cohen's $kappa$ value does. [@cohen1960: @gwet2008; @wongpakaran2013] Given the small number of included studies in this review as a proportion of the total number screened, this is a useful characteristic. The formulae used to calculate the coefficient are given in Appendix \@ref(appendix-gwet).

@@

<!----------------------------------------------------------------------------->
(ref:gwet-caption) Suggested ranges to aid in interpretation of Gwet's AC1 inter-rater reliability metric[@mchugh2012]

@@

@@

#### Random-effects meta-analysis {#meta-analysis-methods}

<!-- TODO May need to talk about FE models here, as I refer to this section for a broader discussion of meta-analysis in the IPD chapter -->

@@

Several of the included studies presented data on multiple categories of lipid levels, but provided an overall effect estimate based on a comparison of only two of these categories (e.g. for example, highest vs lowest quartile). <!-- TODO Question: if I look at highest vs lowest, and also at top dose vs lowest dose, am I double counting results? --> While this allows for easy interpretation of the resulting effect estimate, it ignores any potential non-linear relationships between the exposure and outcome, in addition to discarding useful information contained in the interim groups. In order to address this limitation, I performed a dose-response meta-analysis in those studies reporting more than two categories of lipid levels.

### 2021-12-08 05-Chp-CPRD-Analysis.Rmd
Deal with abbreviations, fix cross-links and some other small edits

Previously known as the General Practice Research Database (GPRD), the Clinical Practice Research Datalink (CPRD) is a large population-based electronic health record (EHR) database.[@herrett2015] The database has been collecting primary care data from participating practices across England since 1987.[@williams2012; @wood2001revitalizing] It contains the primary care records for more than 10 million primary care patients in England, and is broadly representative of the UK population in terms of age, sex and ethnicity.[@herrett2015; @mathur2014]

@@

I considered seven lipid-regulating drug classes based on groupings in the British National Formulary (BNF),[@wishart2017] namely: statins, fibrates, bile acid sequestrants, ezetimibe, nicotinic acid groups, ezetimibe and statin (representing one treatment containing both drugs, rather than the two classes being prescribed concurrently), and omega-3 fatty acid groups.

@@

This published analysis used a propensity matching approach to estimate the association of statins with a range of outcomes in The Health Improvement Network (THIN) database, an alternative source of English electronic health records which has substantial overlap with the CPRD.[@carbonari2015] The code lists used in this analysis were obtained through correspondence with the authors of that study, and are available for inspection (see Section \@ref(cprd-data-avail)).

@@

<!----------------------------------------------------------------------------->
(ref:unadjustedComparisonFig-cap) Results of three models adjusting for a different set of covariates. Note that the x-axis cutoffs (.5,2) are different compared to other plots (0.3,3) to enable greater comparison between the different models.

>>>>

Previously known as the General Practice Research Database, the Clinical Practice Research Datalink (CPRD) is a large population-based EHR database.[@herrett2015] The database has been collecting primary care data from participating practices across England since 1987.[@williams2012; @wood2001revitalizing] It contains the primary care records for more than 10 million primary care patients in England, and is broadly representative of the UK population in terms of age, sex and ethnicity.[@herrett2015; @mathur2014]

@@

I considered seven lipid-regulating drug classes based on groupings in the British National Formulary,[@wishart2017] namely: statins, fibrates, bile acid sequestrants, ezetimibe, nicotinic acid groups, ezetimibe and statin (representing one treatment containing both drugs, rather than the two classes being prescribed concurrently), and omega-3 fatty acid groups.

@@

This published analysis used a propensity matching approach to estimate the association of statins with a range of outcomes in The Health Improvement Network database, an alternative source of English electronic health records which has substantial overlap with the CPRD.[@carbonari2015] The code lists used in this analysis were obtained through correspondence with the authors of that study, and are available for inspection (see Section \@ref(cprd-data-avail)).

@@

<!----------------------------------------------------------------------------->
(ref:unadjustedComparisonFig-cap) Results of three models adjusting for a different set of covariates. Note that the x-axis cutoffs (0.5,2) are different compared to other plots (0.3,3) to enable greater comparison between the different models.

### 2021-12-08 06-Chp-IPD-Analysis.Rmd
Deal with abbreviations, fix cross-links and some other small edits

The first approach focused on previously analysed observational prospective cohort studies examining the effect of blood lipid levels on dementia outcomes, as identified by the systematic review presented in Chapters \@ref(sys-rev-methods-heading)/\@ref(sys-rev-results-heading). The data sources used in each of these cohort analyses were screened against the criteria listed in the previous section, and eligible cohorts were approached for data access. In the first instance, the first/corresponding author of the publication was emailed (see Appendix \@ref(ipd-email-collab) for a copy of the text and documentation sent to potential collaborators) in Autumn 2020. If this approach did not elicit a response within two months, the last author was contacted, on the basis that the first/corresponding author may have been a more junior member of the research group who had moved to a different institution.

@@

@@

Based on available data across cohorts, smoking was classified as never/ever/current, while alcohol consumption was classified as never/ever. Education was categorised into 4 levels (None,O-levels,A-levels,Degree). BMI was treated as a continuous variable while presence of vascular co-morbidities was treated as dichotomous.

@@

<!----------------------------------------------------------------------------->
(ref:mainEffectDem-cap) IPD meta-analysis (model 1) of all-cause dementia outcomes, stratified by lipid fraction

@@

<!----------------------------------------------------------------------------->
(ref:mainEffectVad-cap) IPD meta-analysis (model 1) of vascular dementia outcomes, stratified by lipid fraction. Note that the vascular dementia outcome was only available in the CaPS and Whitehall II cohorts.

>>>>

The first approach focused on previously analysed observational prospective cohort studies examining the effect of blood lipid levels on dementia outcomes, as identified by the systematic review presented in Chapters \@ref(sys-rev-methods-heading)/\@ref(sys-rev-results-heading). The data sources used in each of these cohort analyses were screened against the criteria listed in the previous section, and eligible cohorts were approached for data access. In the first instance, the first/corresponding author of the publication was emailed (see Appendix \@ref(ipd-email-collab) for a copy of the correspondence sent to potential collaborators) in Autumn 2020. If this approach did not elicit a response within two months, the last author was contacted, on the basis that the first/corresponding author may have been a more junior member of the research group who had moved to a different institution.

@@

<!-- TODO May need to rethink this, as introducing model based on missing data which isn't fully discussed until later. -->

@@

Based on available data across cohorts, smoking was classified as never/ever/current, while alcohol consumption was classified as never/ever. Education was categorised into 4 levels (None, O-levels, A-levels, Degree). BMI was treated as a continuous variable while presence of vascular co-morbidities was treated as dichotomous.

@@

<!----------------------------------------------------------------------------->
(ref:mainEffectDem-cap) IPD meta-analysis (Model 1) of all-cause dementia outcomes, stratified by lipid fraction

@@

<!----------------------------------------------------------------------------->
(ref:mainEffectVad-cap) IPD meta-analysis (Model 1) of vascular dementia outcomes, stratified by lipid fraction. Note that the vascular dementia outcome was only available in the CaPS and Whitehall II cohorts.

### 2021-12-08 08-Discussion.Rmd
Deal with abbreviations, fix cross-links and some other small edits

It is customary at this point in the discussion of results based (primarily) on observational data to recommend that a large-scale randomised controlled trial be performed to assess the effect of the proposed intervention. Given the costs and logistical challenges associated with trials examining outcomes with long prodomal periods such as dementia, a more efficient approach is offered by the opportunistic post-trial follow-up (PTFU) of existing RCTs of lipid-regulating agents.

>>>>

It is customary at this point in the discussion of results based (primarily) on observational data to recommend that a large-scale randomised controlled trial be performed to assess the effect of the proposed intervention. Given the costs and logistical challenges associated with trials examining outcomes with long prodomal periods such as dementia, a more efficient approach is offered by the opportunistic post-trial follow-up of existing RCTs of lipid-regulating agents.
